### 2016-11-21 src/group-2-approximation.tex
outline

In this subsection we give a 2-approximation algorithm for general capacities.