### 2016-12-05 src/hardness.tex
working...

Formally, the (unweighted) minimum dominating set problem is defined as
$\min_{D}(|D|)$,
where $D$ is any feasible dominating set.
It is also easy to see that the (unweighted, undirected, uncapacitated) 
maximum carpool matching problem can be defined as $max_{D}(|V \setminus D|)$, 
where $D$ is any feasible dominating set.
The \textsc{Dominating Set-B} is the problem of finding minimum dominating set on 
graphs with degree bounded by $B$,
and it was shown to be APX-hard (for $B \geq 3$) by 
Papadimitriou et al.~\cite{papadimitriou1988optimization}.

Let $G$ be a b-bounded degree graph,
let $D^*(G)$ be a minimum dominating set in $G$,
and let $C^*(G)$ be a maximum carpool matching in $G$, 
then the following lemmas hold:

\begin{lemma}
$|C^*| = |V| - |D^*|$
\end{lemma}

\begin{proof}
The proof follows immediately from the definitions of the two problems. 
\end{proof}

\begin{lemma}
\label{lm:optc_leq_boptd}
$|C^*| \leq b \cdot |D^*|$
\end{lemma}

\begin{proof}
The proof follows directly from the observation that $|D^*| \geq \frac{|V|}{b}$.
\end{proof}

Let $C(G)$ be any feasible carpool matching on $G$, then the following lemma holds:

\begin{lemma}
\label{lm:erreq}
$|C^*| - |C| = (|V| - |C|) - |D^*|$
\end{lemma}

\begin{proof}
$$
|C^*| - |C| 				= 
|V| - |D^*| - |C| 	= 
(|V| - |C|) - |D^*|
$$
\end{proof}

Let $D(G)$ be any feasible dominating set in $G$, 
we argue that $f(G) = G$, and $g(C(G)) = V \setminus C(G)$ 
define a linear reduction (L-reduction) from the \textsc{Dominating Set-B} problem to the 
\textsc{\CARPOOL{}} problem.
\begin{proof}
By lemma~\ref{lm:optc_leq_boptd} we get that $C^*(f(G)) = C^*(G) \leq b \cdot D^*(G)$,
and by lemma~\ref{lm:erreq} we get that 
\begin{equation}
\begin{split}
|C^*(f(G))| - |C(G)|		& = |C^*(G)| - |C(G)| 			\\
							& = (|V| - |C(G)|) - |D^*(G)|	\\
							& = |g(C(G))| - |D^*(G)|
\end{split}
\end{equation}
\end{proof}

\begin{theorem}
The \textsc{\CARPOOL{}} problem is APX-hard.
\end{theorem}

\begin{proof}
The proof follows directly from the above linear reduction and the fact that 
\textsc{Dominating Set-B} is APX-hard. 
\end{proof}