### 2016-12-05 src/main.tex

\documentclass[draft]{article}

\input{in/libs}
\input{in/def}
\input{in/tikz}

\input{in/title}
\input{in/authors}

\begin{document}
\maketitle

\input{outline}

\bibliographystyle{plain}
\bibliography{main}
% \resizebox{.9\textwidth}{!}{\input{tmp}}

\end{document}