### 2016-12-05 src/sub.tex

In this section we show that the problem can be formulated as unconstrained
submodular maximization problem, and thus, it has a 2-approximation algorithm.

Consider the function $w : 2^V \to \R$ where $w(D)$ is the value of the optimal
carpool matching when the set of drivers is $D$.
As we mention in the introduction, $w$ can be computed efficiently.

\begin{theorem}
$w$ is submodular.
\end{theorem}

\begin{proof}
Consider two subsets $A,B \subseteq V$, 
we need to show that $w(A) + w(B) \geq w(A \cup B) + w(A \cap B)$.
Let $E_{A \setminus B} = (V \setminus A) \times (A \setminus B)$, 
$E_{B \setminus A} = (V \setminus A) \times (B \setminus A)$,
and $E_{B \cap A} = (V \setminus (A \cap B)) \times (B \cap A)$.

Let $M_{A \cup B}$ and $M_{A \cap B}$ be the set of edges in an optimal
matching when the set of drivers are $A \cup B$ and $A \cap B$ respectively.
We argue that we can construct two (feasible) matchings $M_A$ and $M_B$ using
the sets of drivers $A$ and $B$ respectively and 
$w(M_A) + w(M_B) = w(M_{A \cup B}) + w(M_{A \cap B})$.
First, take all the edges from $M_{A \cup B} \cap E_{A \setminus B}$ and 
add them to $M_A$, 
add $M_{A \cup B} \cap E_{B \setminus A}$ to $M_B$.
Clearly this can be done.
Now, split the edges from $M_{A \cap B} \cup (M_{A \cup B} \cap E_{A \cap B})$
between $M_A$ and $M_B$ in a way that does not violate the capacity constraints.
One can verify that this can be done.
\end{proof}