### 2016-12-16 src/in/tikz.tex
fig local tight

\usepackage{tikz}
\usetikzlibrary{
	arrows,
	arrows.meta,
	calc,
	graphs,
	patterns,
	positioning,
	shapes,
	decorations.pathmorphing,
	math,
}

%DEFAULT STYLE
\tikzset{default style/.style={
	>=Stealth, 
	on grid, 
	auto, 
	thick,
}}

% NODE
\tikzset{default node/.style={
	draw, 
	circle,
	inner sep=0mm,
	minimum size=5mm,
	very thick,
	font=\small,
	black!70,
}}

% LABELS
\tikzset{label/.style={
	draw=none,
	sloped,
}}

\tikzset{label above/.style={
	label,
	midway,
	above=-1mm,
}}

\tikzset{label below/.style={
	label,
	midway,
	below=-1mm,
}}