### 2016-12-19 src/abstract.tex
intro

The \textsc{\CARPOOL{}} problem is a star packing problem in directed graphs.
Formally, given a directed graph $G = (V, A)$,
a capacity function $c: V \rightarrow \N$,
and a weight function $w : A \rightarrow \R $,
a feasible \emph{carpool matching} is a subset of arcs, 
$M \subseteq A$, where
$\forall v \in V, 
\din[M](v) \cdot \dout[M](v) = 0, 
\din[M] \leq c(v),
\dout[M] \leq 1$. 
In the \textsc{\CARPOOL{}} problem we seek for a matching $M$
that maximizes the total weight of $M$.

The problem arises when designing an online carpool service, 
such as Zimride~\cite{zimride}, 
that tries to connect between passengers and drivers based on (arbitrary) similarity function.
The problem is known to be NP-hard, 
even for uniform weights and without capacity constraints.

we show that the problem can be formulated as
unconstrained submodular maximization problem, 
thus it has a $\frac{1}{2}$-approximation algorithm.
For the unweighted variant of the problem when the maximum possible capacity, $C$, is
fixed, we give a $(\frac{C + 1}{2C} - \varepsilon)$-approximation algorithm
for the any $\varepsilon > 0$.

We also consider a variant of the \textsc{\CARPOOL{}} problem where every arc
has also an integral size $s:A \to \N$ and the constraint $\din[M] \leq c(v)$
is replaced with the constraint $\sum_{(u,v)}s(u,v) \leq c(v)$.
We show that this variant of the problem can not be formulated as an
unconstrained submodular maximization as before.
We give, however, a ($2 + \varepsilon$)-approximation algorithm for this
variant, and argue that the local search algorithm can be generalized to this
variant of the problem when $C$ is fixed.