### 2017-01-30 src/abstract.tex
Merge Dror

The \carpool problem is a star packing problem in directed graphs.
Formally, given a directed graph $G = (V, A)$, a capacity function $c:
V \rightarrow \N$, and a weight function $w : A \rightarrow \R $, a
feasible \emph{carpool matching} is a subset of arcs, $M \subseteq A$,
such that every $v \in V$ satisfies:%
\begin{inparaenum}[(i)]
\item $\din[M](v) \cdot \dout[M](v) = 0$,
\item $\din[M](v) \leq c(v)$, and 
\item $\dout[M](v) \leq 1$.
\end{inparaenum}
A vertex $v$ for which $\dout[M](v) = 1$ is a \emph{passenger}, and a
vertex for which $\dout[M](v) = 0$ is a driver who has $\din[M](v)$
passengers.  In the \carpool problem the goal is to find a matching
$M$ of maximum total weight.

The problem arises when designing an online carpool service, such as
Zimride~\cite{zimride}, which tries to connect between passengers and
drivers based on (arbitrary) similarity function.  The problem is
known to be NP-hard, even for uniform weights and infinite capacities.
%without capacity constraints.

We show that the problem can be formulated as an unconstrained
submodular maximization problem, thus it admits a
$\frac{1}{2}$-approximation algorithm.  For the unweighted variant of
the problem when the maximum possible capacity, $C$, is fixed, we give
a $(\frac{C + 1}{2C} - \varepsilon)$-approximation algorithm for the
any $\varepsilon > 0$.

We also show that the
problem remains APX-hard even if $C \leq 3$.

We also consider an extension, called \gcp, in which each vertex
represents a group of passengers.  Group sizes are non-uniform, and we
consider the case where a group size may change as a function of the
driver.  Formally, each arc $(u,v) \in A$ has a size $s(u,v) \in \N$,
and the constraint $\din[M](v) \leq c(v)$ is replaced with
$\sum_{u:(u,v) \in M} s(u,v) \leq c(v)$.  We show that \gcp cannot be
formulated as an unconstrained submodular maximization problem.
Nevertheless, we give a $(\frac{1}{2} - \varepsilon)$-approximation
algorithm for this variant, and argue that the local search algorithm
can be generalized to this variant of the problem when $C$ is fixed.