### 2017-01-30 src/main.tex

%\documentclass[11pt,draft]{article}
%\documentclass[a4paper,11pt]{article}
\documentclass[a4paper,UKenglish]{lipics-v2016}

\usepackage{microtype}%if unwanted, comment out or use option "draft"

\usepackage{amsmath,amssymb,mathtools,amsthm}
\usepackage[linesnumbered, ruled]{algorithm2e}
\usepackage{xspace}
\usepackage{subfig}
\usepackage{paralist}
%\usepackage{hyperref}
%\hypersetup{final}
%\usepackage[inline]{enumitem}

% DEBUG
%\usepackage{fullpage}
\usepackage{todonotes}
%\usepackage{lineno}
%\linenumbers

% EXTRA
%\usepackage{authblk}

\input{in/def}
\input{in/tikz}

\renewcommand{\floatpagefraction}{0.7}

\sloppy

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Improved Approximation Algorithms for the
Maximum Carpool Matching Problem}

\titlerunning{Improved Approximation Algorithms for Maximum Carpool Matching}

\author[1]{Gilad Kutiel}
\author[2]{Dror Rawitz%
\thanks{Supported in part by the Israel Science Foundation (grant no.~497/14).}}

\affil[1]{Department of Computer Science, Technion, Haifa,
  Israel. \\ \texttt{gkutiel@cs.technion.ac.il}}
\affil[2]{Faculty of Engineering, Bar Ilan University, Ramat Gan,
  Israel. \\ \texttt{dror.rawitz@biu.ac.il}}

\authorrunning{G. Kutiel and D. Rawitz}

\Copyright{G. Kutiel and D. Rawitz}

\subjclass{F.2.2 Nonnumerical Algorithms and Problems}

\keywords{approximation algorithms, local search}

%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\EventEditors{John Q. Open and Joan R. Acces}
\EventNoEds{2}
\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
\EventShortTitle{CVIT 2016}
\EventAcronym{CVIT}
\EventYear{2016}
\EventDate{December 24--27, 2016}
\EventLocation{Little Whinging, United Kingdom}
\EventLogo{}
\SeriesVolume{42}
\ArticleNo{23}
% Editor-only macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\maketitle

\input{outline}

\subparagraph*{Acknowledgements.}

We thank David Adjiashvili and Reuven Bar-Yehuda for helpful
discussions.

%\bibliographystyle{plain}
\bibliographystyle{plainurl}% the recommended bibstyle
\bibliography{main}

\end{document}