### 2017-01-30 src/preliminaries.tex

In this section

We use standard graph theoretic notation.

$N^-$ and $N^+$.

Define $\deg$

Define $\deg_M(v) \eqdf \din[M](v) + \dout[M](v)$
%
For a subset $U \subseteq V$ of vertices define
$\deg_M(U) \eqdf \sum_{v \in U} \deg_M(u)$.

We define $c_{\max} \eqdf \max_{v \in V} c(v)$.

and $w_{\max} \eqdf \max_{(v,u) \in E} w(v)$.

Given a subset $A' \subseteq A$, we define
$w(A') \eqdf \sum_{(v,u) \in A'} w(v,u)$.

Given a directed graph $G = (V,A)$ and a feasible carpool matching
$M \subseteq A$, we define a partition of $V$ as follows:
\begin{align*}
P_M & \eqdf \set{v : \dout[M](v) = 1} \\
D_M & \eqdf \set{v : \din[M](v) \geq 1} \\
Z_M & \eqdf \set{v : \dout[M](v) = \din[M](c) = 0}
~.
\end{align*}
where $P_M$ is the set of \emph{passengers}, $D_M$ is the set of
\emph{drivers}, and $Z_M$ is the set of \emph{singletons}.

In Figure~\ref{fig: }