### 2017-01-31 src/in/title.tex

\title{Improved Approximation Algorithms for the
Maximum Carpool Matching Problem}

\titlerunning{Improved Approximation Algorithms for Maximum Carpool Matching}

\author[1]{Gilad Kutiel}
\author[2]{Dror Rawitz%
\thanks{Supported in part by the Israel Science Foundation (grant no.~497/14).}}

\affil[1]{Department of Computer Science, Technion, Haifa,
  Israel. \\ \texttt{gkutiel@cs.technion.ac.il}}
\affil[2]{Faculty of Engineering, Bar Ilan University, Ramat Gan,
  Israel. \\ \texttt{dror.rawitz@biu.ac.il}}

\authorrunning{G. Kutiel and D. Rawitz}

\Copyright{G. Kutiel and D. Rawitz}

\subjclass{F.2.2 Nonnumerical Algorithms and Problems}

\keywords{approximation algorithms, local search}