### 2017-01-31 src/main.tex

%\documentclass[11pt,draft]{article}
%\documentclass[a4paper,11pt]{article}
\documentclass[a4paper,UKenglish]{lipics-v2016}

\usepackage{microtype}%if unwanted, comment out or use option "draft"

\usepackage{amsmath,amssymb,mathtools,amsthm}
\usepackage[linesnumbered, ruled]{algorithm2e}
\usepackage{xspace}
\usepackage{subcaption}
\usepackage{paralist}

% DEBUG
\usepackage{todonotes}

\input{in/def}
\input{in/tikz}

\renewcommand{\floatpagefraction}{0.7}

\sloppy

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{in/title}
%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\EventEditors{John Q. Open and Joan R. Acces}
\EventNoEds{2}
\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
\EventShortTitle{CVIT 2016}
\EventAcronym{CVIT}
\EventYear{2016}
\EventDate{December 24--27, 2016}
\EventLocation{Little Whinging, United Kingdom}
\EventLogo{}
\SeriesVolume{42}
\ArticleNo{23}
% Editor-only macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\maketitle

\input{outline}

\subparagraph*{Acknowledgements.}

We thank David Adjiashvili and Reuven Bar-Yehuda for helpful
discussions.

\bibliographystyle{plainurl}
\bibliography{main}

\end{document}