### 2017-02-09 src/preliminaries.tex

$\nin[]$ and $\nout[]$.

Define $\deg$

Define $\deg_M(v) \eqdf \din[M](v) + \dout[M](v)$
%
For a subset $U \subseteq V$ of vertices define
$\deg_M(U) \eqdf \sum_{v \in U} \deg_M(u)$.

We define $c_{\max} \eqdf \max_{v \in V} c(v)$.

and $w_{\max} \eqdf \max_{(v,u) \in E} w(v)$.

Given a subset $A' \subseteq A$, we define
$w(A') \eqdf \sum_{(v,u) \in A'} w(v,u)$.