### 2017-02-09 src/in/title.tex
refactor

\title{Improved Approximation Algorithms for the
Maximum Carpool Matching Problem}

\titlerunning{Improved Approximation Algorithms for Maximum Carpool Matching}

\author[1]{Gilad Kutiel}
\author[2]{Dror Rawitz%
\thanks{Supported in part by the Israel Science Foundation (grant no.~497/14).}}

\affil[1]{Department of Computer Science, Technion, Haifa,
  Israel. \\ \texttt{gkutiel@cs.technion.ac.il}}
\affil[2]{Faculty of Engineering, Bar Ilan University, Ramat Gan,
  Israel. \\ \texttt{dror.rawitz@biu.ac.il}}

\authorrunning{G. Kutiel and D. Rawitz}

\Copyright{G. Kutiel and D. Rawitz}

\subjclass{F.2.2 Nonnumerical Algorithms and Problems}

\keywords{approximation algorithms, local search}

%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\EventEditors{John Q. Open and Joan R. Acces}
\EventNoEds{2}
\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
\EventShortTitle{CVIT 2016}
\EventAcronym{CVIT}
\EventYear{2016}
\EventDate{December 24--27, 2016}
\EventLocation{Little Whinging, United Kingdom}
\EventLogo{}
\SeriesVolume{42}
\ArticleNo{23}
% Editor-only macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%