### 2017-02-09 src/main.tex

%\documentclass[11pt,draft]{article}
%\documentclass[a4paper,11pt]{article}
\documentclass[a4paper,UKenglish]{lipics-v2016}

\input{in/def}
\input{in/tikz}
\input{in/libs}

\renewcommand{\floatpagefraction}{0.7}

\sloppy

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{in/title}

\begin{document}

\maketitle

\input{outline}

\subparagraph*{Acknowledgements.}

We thank David Adjiashvili and Reuven Bar-Yehuda for helpful
discussions.

\bibliographystyle{plainurl}
\bibliography{main}

\end{document}