### 2017-02-09 src/group-2-approximation.tex
rm todos

In this subsection we give a $\frac{1}{2}$-approximation algorithm for general
capacities.
First, let us introduce a $(\frac{1}{2} - \varepsilon)$-approximation algorithm
for the original problem.
We then show how to generalize the algorithm to work also for the group carpool
matching.

\tikzset{m/.style={
	blue, 
	dashed,	
}}

Let $M$ be a feasible carpool matching,
the \emph{value} of a vertex $v$, denoted $val(v)$, 
is the sum of the weights of the arcs in $M$ that incident $v$. 
That is: 
%
\begin{definition}[$val$]
$val_M(v) = \sum_{(u, v) \in M} w(u, v) + \sum_{(v, u) \in M} w(v, u)$
\end{definition}
%
If $V$ is a set of vertices, then $val_M(V) \triangleq \sum_{v \in V} val_M(v)$.
%
Denote by $\delta(e)$ the difference between the weight of the arc and the value
of its source vertex,
that is:
%
\begin{definition}[$\delta$]
$\delta_M(u, v) = w(u, v) - val_M(u)$
\end{definition}
%
If $S$ is a set of arcs, then $\delta(S) \triangleq \sum_{e \in S}\delta(v)$.

An \emph{improvement} to vertex $v$ 
is a set of arcs entering $v$ of size not greater than the capacity of $v$, 
for which the sum of the $\delta$s is greater than the value of $v$,
formally:

\begin{definition}[improvement]
$S_v \subseteq V \times \{v\} \cap A$ is an \emph{improvement} (w.r.t $M$)
if the following conditions hold:
\begin{itemize}
\item
$|S_v| \leq c(v)$
\item
$\delta_M(S_v) > val_M(v)$
\end{itemize}  
\end{definition}
%
If, for vertex $v$, an improvement exists, we say that vertex $v$ can be \emph{improved}.
%
Given an arc $(u,v)$, 
let $inc(u,v)$ be the set of arcs that incident $(u, v)$,
formally:
%
\begin{definition}[inc]
$inc(u,v) = (V \times \{u, v\} \cup \{u, v\} \times V) \cap A$
\end{definition}
%
If $S$ is a set of arcs, then $inc(S) \triangleq \bigcup_{e \in S} inc(e)$.
%
The \emph{passengers} of $v$ (w.r.t some set of arcs) are all the vertices that have
outgoing arc to $v$, formally: 
%
\begin{definition}[passengers]
$\text{pass}_M(v) = \{u : (u, v) \in M\}$
\end{definition}
%
Figure~\ref{fig:defs} depicts all the above definitions.

\begin{figure}
\centering
\input{fig-val-delta}
\caption[]{
\label{fig:defs}
In this example, $M$ is the set of the blue, dashed arcs, then:
$val_M(2) = 7$,
$val_M(3) = 2$,
$val_M(6) = 0$,
$\delta_M(2, 3) = 1$,
$\delta_M(6, 3) = 2$,
the set $\{(2,3), (6,3)\}$ is an \emph{improvement} to vertex 3,
$inc(2,3) = \{(1,2),(4,2),(3,5),(6,3)\}$,
$\text{pass}_M(2) = \{1, 4\}$.
}
\end{figure}

We are now ready to describe our algorithm.
The algorithm starts with an empty matching,
in every iteration it seeks for a vertex that can be improved,
it then improve the matching by adding
the arcs that improve the vertex, 
just after removing the arcs that incident them.
The algorithm halts when no vertex can be improved.
Figure~\ref{fig:improvement} depicts an improvement step,
and the full algorithm is listed in Algorithm~\ref{alg:grd}.

\begin{figure}
\centering
\input{fig-improvement}
\caption[]{
\label{fig:improvement}
\subref{sub:can improved}
A matching that can be improved
\subref{sub:improved}
The matching after improving vertex 3
}
\end{figure}

\begin{algorithm}
\caption{
\label{alg:grd}
GRD
}
$M \leftarrow \emptyset$									\\
\While{done}{
	done $\leftarrow$ True									\\
	\For{$v \in V$}{
		\If{$\exists$~improvment~$S_v$}{
			$M \leftarrow M \setminus inc(S_v) \cup S_v$	\\
			done $\leftarrow$ False							\\
		}
	}
}
\end{algorithm}

\begin{remark}
\label{rem:improve}
Determining if a vertex $v$ can be improved can be done efficiently
by considering the incoming arcs to $v$ in a decreasing order of their $\delta$s,
and only ones with positive value.
Vertex $v$ can be improved, then, if the $\delta$s of the first $c(v)$ (or less) arcs
sum up to more than $val_M(v)$.
\end{remark}

\begin{remark}
It is easy to see, if the weights on the edges are integral and polynomial
bounded the algorithm runs in polynomial time, this is because in each iteration the
algorithm improves the value of the matching by at least one or otherwise it
terminates.
\end{remark}

\begin{remark}
When the weights are not integral nor polynomial bounded one can use scaling and
rounding first to ensure polynomial runtime in the cost of $\varepsilon$ in the
approximation ratio.
\end{remark}

We now argue that, with respect to any matching, 
the total value of all the vertices equals twice the value of the matching, 
formally: 
\begin{lemma}
\label{lm:val-twice}
$\sum_{v \in V} val_M(v) = 2 \sum_{e \in M} w(e)$
\end{lemma}

\begin{proof}
\begin{equation}
\begin{split}
\sum_{v \in V} val_M(v)	& = 
\sum_{v \in V} \left( \sum_{(u, v) \in M} w(u, v) + \sum_{(v, u) \in M} w(v, u) \right)	\\
						& = \sum_{v \in V}\sum_{(u, v) \in M} w(u, v) + 
							\sum_{v \in V}\sum_{(v, u) \in M} w(v, u)					\\
						& = 2 \sum_{e \in M} w(e)
\end{split}
\end{equation}
\end{proof}

Let $M$ be a matching, and let $v$ be a vertex with no improvement,
let $S \subseteq V \times \{v\} \cap A$ of size not greater than $c(v)$,
then the following lemma holds:

\begin{lemma}
\label{lm:no improve}
$w(S) \leq val_M(v) + val_M(\text{pass}_S(v))$
\end{lemma}

\begin{proof}
If no improvement exists then
$$
w(S) - val_M(\text{pass}_S(v))=
\sum_{(u,v) \in S}(w(u,v) - val_M(u)) =
\delta_M(S) 
\leq val_M(v)
$$
\end{proof}

To bound the approximation ratio of the GRD algorithm, 
we use a charging scheme argument.
Given the matching produced by the algorithm, 
we load every vertex with amount of money equal to its value,
we then show that this is enough to pay for every arc in an optimal matching.

\begin{theorem}
Algorithm~\ref{alg:grd} is $\frac{1}{2}$-approximation.
\end{theorem}

\begin{proof}
Let $M$ be the matching produced by algorithm~\ref{alg:grd}, 
and let $M^*$ be an optimal matching.
Load every vertex with $val_M(v)$ money, 
by lemma~\ref{lm:val-twice} this is exactly twice the value of $M$.
Consider a driver $v$ in $M^*$, and let $S = V \times \{v\} \cap M^*$.
By lemma~\ref{lm:no improve} we know that $w(S) \leq val_M(v) + val_M(\text{pass}_S(v))$,
thus we can pay for $S$, using the money on $v$ and $\text{pass}_S(v)$.
Clearly, these vertices will not be charged again.
\end{proof}

We now argue that this algorithm can be generalized to solve the group carpool
matching problem.
The main concern when trying to adopt the algorithm to the group carpool
matching is how to determine if a vertex can be improved.
Observe that Remark~\ref{rem:improve} does not hold anymore.
If the maximum possible capacity is bounded, one can test for improvement of
$v$ by considering all the subsets of edges intersecting $v$. 
When the capacity is unbounded, however, it is easy to see that efficient
way to test improvement implies a solution to the Knapsack problem.
Indeed, if this is the case, one can test for improvement by using a known FPTAS
for Knapsack, this is in the cost of additional $\varepsilon$ reduction in the
approximation ratio of the algorithm.

To conclude this section we show that our analysis is tight, 
as depicts in Figure~\ref{fig:grd worst}.
%\todo[inline]{what happen if we apply the rule on both directions ?}
\begin{figure}
\centering
\input{fig-2-tight}
\caption[]{
\label{fig:grd worst}
GRD algorithm, worst case example: \\
Consider a path with $2n + 1$ arcs,
and alternating arc weights (2 and 1),
if the GRD algorithm selects all the 1 weighted arcs,
then no further improvement can be done and the value of the matching is $n + 1$,
while the optimal matching has value of $2n$.
}
\end{figure}