### 2017-02-09 src/group-sub.tex

Consider the submodular maximization formulation given in Section~\ref{sec:sub}.
Figure~\ref{fig:not submodular} shows that the function $w : 2^V \to \R$ is not
submodular anymore.

\begin{figure}
\begin{center}
\input{fig-not-submodular}
\end{center}
\caption{Consider the following instance where node capacity are
written inside nodes, arcs are labeled with their size and all weights
are equal to 1.  It is easy to see that if $f(S)$ is the optimal
solution where $S$ is the set of drivers, then $f(A) + f(B) = 2 \leq 3
= f(A \cup B) + f(A \cap B)$, thus, $f$ is not submodular.}
\label{fig:not submodular}
\end{figure}