### 2017-02-09 src/group.tex

We now consider a variant of the problem when, in addition to the
capacity function $c:V \to \N$, and the weight (profit) function
$w:A \to \R$, we also have a size function $s:V \to \N$, and the
constraint $\din[M] \leq c(v)$ is replaced with the constraint
$\sum_{(u,v)}s(u) \leq c(v)$.  The motivation for this variant of
the problem is when we want to allow a group of people to travel
together in the same car, thus $s$ is the size of the group.  We start
by showing that this variant of the problem does not fit the
submodular maximization formulation anymore. We show, however, that
the problem still has a $(2 +\eps)$-approximation algorithm.  Finally,
we argue that the local search algorithm can be trivially generalized
to handle the new variant of the problem while keeping the same
approximation guarantees.