### 2017-02-10 src/group-sub.tex

Consider the submodular maximization formulation given in
Section~\ref{sec:sub}, namely $\barw: 2^V \to \R$, where
$\barw(S) \eqdf w(M(S))$ and $M(S)$ is the maximum weight carpool
matching that satisfies $D_{M(S)} \subseteq S \subseteq V \setminus
P_{M(S)}$.
%
Figure~\ref{fig:not submodular} contains an instance that shows that
the function $\barw$ is not submodular anymore.

\begin{figure}
\begin{center}
\input{fig-not-submodular}
\end{center}
\caption{Consider the following instance where node capacity are
written inside nodes, arcs are labeled with their size and all weights
are equal to 1.  We have that $f(A) + f(B) = 2 < 3 = f(A \cup B) +
f(A \cap B)$.}
\label{fig:not submodular}
\end{figure}