### 2017-02-10 src/group.tex

We now consider \gcp which a variant of \carpool in which in addition
to the capacity function $c:V \to \N$, and the weight function
$w:A \to \R^+$, we also have a size function $s:V \to \N$, and the
constraint $\din[M](v) \leq c(v)$ is replaced with the constraint
$\sum_{u:(u,v) \in M} s(u) \leq c(v)$.
%
%The motivation for this variant of the problem is when we want to
%allow a group of people to travel together in the same car, thus $s$
%is the size of the group.
%
We start by showing that this variant of the problem does not fit the
submodular maximization formulation anymore.  We show, however, that
the problem still has a $(\half -\eps)$-approximation algorithm by
extending the local improvement algorithm from
Section~\ref{sec:improve}.
%
Finally, we show that the local search algorithm from
Section~\ref{sec:local} can be generalized to handle \gcp while
keeping the same approximation guarantees.