### 2017-02-10 src/main.tex

\documentclass[a4paper,UKenglish]{lipics-v2016}

\input{in/def}
\input{in/tikz}
\input{in/libs}

\renewcommand{\floatpagefraction}{0.8}

\sloppy

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{in/title}

\begin{document}

\maketitle

\input{outline}

\subparagraph*{Acknowledgements.}

We thank David Adjiashvili and Reuven Bar-Yehuda for helpful
discussions.

\newpage

\bibliographystyle{plainurl}
\bibliography{main}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage

\appendix

\input{appendix}

\end{document}