### 2017-02-13 src/in/tikz.tex

\usepackage{tikz}
\usetikzlibrary{
	arrows.meta,
	calc,
	graphs,
	patterns,
	positioning,
	shapes,
	decorations.pathmorphing,
	math,
}

% NODE
\tikzset{default node/.style={draw,circle,inner sep=0mm,minimum size=5mm,very thick,font=\small,black!70,}}

% LABELS
\tikzset{label/.style={draw=none,sloped,}}
\tikzset{label above/.style={label,midway,above=-1mm,}}
\tikzset{label below/.style={label,midway,below=-1mm,}}
\tikzset{outline/.style={loosely dashed, ultra thin, lightgray}}
\tikzset{m/.style={blue, dashed}}

%DEFAULT STYLE
\tikzset{
	>=Triangle[], 
	on grid, 
	auto, 
	thick,
}