### 2017-02-15 src/in/title.tex

\title{%Improved Approximation
Local Search Algorithms for the Maximum Carpool Matching Problem}

\titlerunning{%Improved
Local Search Approximation Algorithms for Maximum Carpool Matching}

\author[1]{Gilad Kutiel}
\author[2]{Dror Rawitz%
\thanks{Supported in part by the Israel Science Foundation (grant no.~497/14).}}

\affil[1]{Department of Computer Science, Technion, Haifa,
  Israel. \\ \texttt{gkutiel@cs.technion.ac.il}}
\affil[2]{Faculty of Engineering, Bar Ilan University, Ramat Gan,
  Israel. \\ \texttt{dror.rawitz@biu.ac.il}}

\authorrunning{G. Kutiel and D. Rawitz}

\Copyright{G. Kutiel and D. Rawitz}

\subjclass{F.2.2 Nonnumerical Algorithms and Problems}

\keywords{approximation algorithms, local search}

%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\EventEditors{John Q. Open and Joan R. Acces}
%\EventNoEds{2}
%\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
%\EventShortTitle{CVIT 2016}
%\EventAcronym{CVIT}
\EventYear{2017}
%\EventDate{December 24--27, 2016}
%\EventLocation{Little Whinging, United Kingdom}
\EventLogo{}
\SeriesVolume{1} %42
\ArticleNo{1} %23
% Editor-only macros::end %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%