### 2017-02-15 src/local.tex

In this section we present a local search $(\half + \inv{2c_{\max}}
- \eps)$-approximation algoithm for unweighted \carpool whose running
time is polynomial if $c_{\max} = O(1)$.

Let $k$ be a constant integer to be determined later.
Algorithm~\textbf{EdgeSwap} (Algorithm~\ref{alg:local}) maintains a
feasible matching $M$ throughout its execution and operates in
iterative manner where in each iteration it tries to find a better
solution by replacing a subset of at most $k$ edges in the current
solution with another (larger) subset of edges not in the solution.
The algorithm halts when no improvement can be done.

\begin{algorithm}
\caption{\textbf{EdgeSwap}$(G,c,k)$}
\label{alg:local}
\begin{small}
\SetKw{True}{true}
\SetKw{False}{false}
%\KwIn{$G = (V, A)$, $c : V \rightarrow \N$, $k$}
%\KwOut{$M$}
$M \leftarrow \emptyset$								\\
\Repeat{done}{
\label{line:outerloop}
 	$done \leftarrow{}$ \True							\\
 	\ForAll{$M' \subseteq M : |M'| \leq k$}{
 		\ForAll{$A' \subseteq A \setminus M : |A'| = |M'| + 1$}{
			\If{$M \setminus M' \cup A'$ is feasible}{
				$M \leftarrow{} M \setminus M' \cup A'$	\\
				$done \leftarrow{}$ \False				\\
			}
		}
 	}
}
\Return{$M$}
\end{small}
\end{algorithm}

Algorithm~\textbf{EdgeSwap} terminates in polynomial time, since in
every non-final iteration it improves the value of the solution by
one.  Thus, after at most $n$ iterations the algorithm terminates.  In
every iteration the algorithm examines all subsets of edges of a fixed
size and tests for feasibility, both these operations can be done in
polynomial time.

Observe that a vertex $v \in D_M \cup Z_M$ is the center of
a \emph{directed star} whose leaves are the passengers in the set
$P_M(v) = \set{u : (u,v) \in M}$
($P_M(v) = \emptyset$, for $v \in Z_M$).
%
Given a carpool matching $M$, we define $\calS(M)$ to be the set of
stars that are induced by $M$.  Denote by $V(S)$ the set of vertices
of a star, i.e., if $v$ is the center of $S$, then $V(S) = \{v\} \cup
P_M(v)$.  Also, let $A(S)$ be the arcs of $S$.  For
$\calT \subseteq \calS(M)$, define
$V(\calT) \eqdf \bigcup_{S \in \calT} V(S)$ and
$A(\calT) \eqdf \bigcup_{S \in \calT} A(S)$.

It remains to analyze the approximation ratio of \textbf{EdgeSwap}.
Consider an optimal matching $M^*$, and $M$ be the matching computed
by \textbf{EdgeSwap}.  Given both matchings we build the \emph{star
graph} of the two solutions in which each node represents a star from
the optimal solution, namely from $\calS(M^*)$, and an edge exists
between two nodes if there is a star in $\calS(M)$ that intersects the
two corresponding stars of the optimal solution.
%
Formally $H = (\calS(M^*), E)$ where
$
{E = \set{(S^*_i, S^*_j) : \exists S \in \calS(M), 
         V(S) \cap V(S^*_i) \neq \emptyset \land
         V(S) \cap V(S^*_j) \neq \emptyset}}
$.
Figure~\ref{fig:conflict} depicts a star graph.

\begin{figure}[t]
%
\begin{subfigure}[t]{.4\linewidth}
\caption{\label{subfloat:graph}}
\input{fig-sub-graph}
\end{subfigure}
%
\hfill
%
\begin{subfigure}[t]{.4\linewidth}
\caption{\label{subfloat:conflict}}
\input{fig-sub-conflict}
\end{subfigure}
%
\caption[]{
\subref{subfloat:graph} 
An optimal matching depicted by the dashed red edges,
and another solution depicted by the solid green edges.
The optimal stars are outlined.  
\subref{subfloat:conflict}
The star graph, each vertex corresponded to a star in the original graph. 
The upper left vertex corresponds to the star that contains the vertices a,b,c.
The numbers inside the vertices corresponds to $d_i$.   
}
\label{fig:conflict}
\end{figure}

\begin{lemma}
The maximum degree of $H$ is $\cmax(\cmax+1)$.
\end{lemma}
\begin{proof}
Each star in $\calS(M^*)$ contains at most $\cmax+1$ vertices and each
such vertex can belong to a star in $\calS(M)$ containing additional
$\cmax$ vertices, each of which is located in a different star in
$\calS(M^*)$.
\end{proof}

In what follows we compare $\abs{M}$ and $\abs{M^*}$ in maximal
connected components of the star graph $H$.  Intuitively, we show that
$M$ is optimal on small maximal components, and that the approximation
ratio on medium (non-necessarily) components can be bounded due to the
termination condition of \textbf{EdgeSwap}.  Large maximal components
will be partitioned into medium components.

We first show that large connected graphs (or maximal connected
components) can be partitioned in to medium size components.  The
proof is given in Appnedix~\ref{sec:omitted}.

\begin{lemma}
\label{lemma:dec}
An undirected connected graph $G = (V,E)$ with maximum degree
$\Delta$, can be decomposed into connected components of size at least
$\ell$ and at most $\Delta \ell$, if $\ell \leq \abs{V}$.
\end{lemma}

\iffalse %%%%%% Removed

\begin{proof}
Let $T$ be any spanning tree of $G$ and let $r$ be an arbitrary root.
Starting with $r$, repeatedly choose a child whose subtree is strictly
larger than $\frac{\Delta-1}{\Delta} \abs{V}$, until this is not
possible.  Hence, we have reached a vertex $v$ whose subtree contains
more than $\frac{\Delta-1}{\Delta} \abs{V}$ vertices, but the subtrees
of its children contain at most $\frac{\Delta-1}{\Delta} \abs{V}$
vertices.
%
If $v = r$, then $v$ has a child $u$ with at least $\inv{\Delta}
(\abs{V}-1) \geq \ell$ vertices.  Otherwise, there exists a child $u$
of $v$ whose subtree is of size at least
$\inv{\Delta} \abs{V} \geq \ell$.  Disconnect $u$ and the vertices in
its subtree from the graph.
%
In both cases $u$'s subtree contains at most
$\frac{\Delta-1}{\Delta} \abs{V}$ vertices, and thus at least
$\inv{\Delta} \abs{V} \geq \ell$ vertices remain in the graph.  Hence
we obtain two connected subgraphs of size at least $\ell$.
%
We repeat this procedure recursively on any component of size
strictly larger $\Delta\ell$.
\end{proof}

\fi %%%%%% End removed

Define $\deg_M(v) \eqdf \din[M](v) + \dout[M](v)$.  For a subset
$U \subseteq V$ of vertices define $\deg_M(U) \eqdf \sum_{v \in
U} \deg_M(u)$.
%
Observe that $\abs{M} = \half \deg_M(V)$.

In the next lemma we bound the degree ratio in a component that
contains stars with at most $k$ arcs.

\begin{lemma}
\label{lemma:r}
Let $\calT \subseteq \calS(M^*)$ that induces a connected subgraph of
$H$.  If $\abs{A(\calT)} \leq k$, then
\[
\frac{\deg_M(V(\calT))}{\deg_{M^*}(V(\calT))} 
\geq \half + \inv{2\cmax} - \inv{2\cmax\abs{\calT}}
~.
\]
\end{lemma}
\begin{proof}
Consider the solution $M'$ obtained from $M$ by removing all the
edges from $M$ that intersect $V(\calT)$ and adding all the edges from
$M^*$ that intersect $V(\calT)$.  Observe that if an edge $(u,v)$ in
$M^*$ intersects $V(\calT)$, then $\set{u,v} \in V(\calT)$ by the
definition of the graph $H$.  Hence, $M'$ is feasible carpool
matching.

Since $\calT$ induces a connected subgraph of $H$, the removal of
edges in $M$ that intersect $V(\calT)$ decreased $\abs{M}$ by at most
$\deg_M(V(\calT)) - \abs{\calT} + 1$.
%
On the other hand, the increase in size is exactly
 $\half \deg_{M^*}(V(\calT)) \leq \cmax \abs{\calT}$.
%
Since $\abs{A(\calT)} \leq k$, we know that this difference can not be
positive, or else, \textsc{EdgeSwap} would not have terminated.  Thus
\(
\half \deg_{M^*}(V(\calT)) \leq \deg_M(V(\calT)) - \abs{\calT} + 1
\),
and so
\[
\frac{\deg_M(V(\calT))}{\deg_{M^*}(V(\calT))}
\geq \half + \frac{\abs{\calT} - 1}{\deg_{M^*}(V(\calT))}
\geq \half + \frac{\abs{\calT} - 1}{2\cmax \abs{\calT}}
=    \half + \inv{2\cmax} - \inv{2\cmax \abs{\calT}}
~,
\]
as required.
\end{proof}

\iffalse %%%%% Removed

\begin{lemma}
%\label{lemma:dec}
Given a parameter $k$, a connected graph $G$ with a maximum degree $d$
can be decomposed into connected components of size at least $k$, and
at most $d(k-1) + 1$.
\end{lemma}

\begin{proof}
We give a constructive proof.  Start constructing a connected
component by adding adjacent vertices one by one to the component and
removing them from the graph.  If at some point, removing a vertex
disconnect the graph, add all the small components (of size less than
$k$) to the constructed component and decompose the large components
(of size at least $k$) recursively.  Note that removing a vertex can
break the graph to at most $d - 1$ additional components.
\end{proof}

\fi %%%%% End removed

It remains to bound the approximation ratio of \textbf{EdgeSwap}.

\begin{lemma}
If $k \geq \cmax$, then $\abs{M} \geq (\half + \inv{2\cmax}
- \frac{\cmax(\cmax+1)}{2k}) \cdot \abs{M^*}$.
\end{lemma}
\begin{proof}
Consider a maximal (with respect to set inclusion) connected component
of $H$ induced by the vertices in $\calT \subseteq \calS(M^*)$.
%
If $\abs{M \cap A(\calT)} \leq k$, then it must be that $\abs{M \cap
A(\calT)} = \abs{M^* \cap A(\calT)}$, since otherwise $M \cap
A(\calT)$ could be improved.

It remains to consider a maximal component $\calT$ such that If
$\abs{M \cap A(\calT)} > k$.  Since the number of edges in
$S \in \calS(M^*)$ is at most $\cmax$, it must be that $\abs{V(\calT)}
> \frac{k}{\cmax}$.
%
Due to Lemma~\ref{lemma:dec} (with $\ell
= \frac{k}{\cmax^2(\cmax+1)}$) we can partition $\calT$ into connected
components each of which contains between $\frac{k}{\cmax^2(\cmax+1)}$
and $\frac{k}{\cmax}$ vertices.
%
Since each such vertex set $\calX$ is connected and contains at most
$\frac{k}{\cmax}$ stars, it follows that $\abs{A(\calX)} \leq k$.
%
Due to Lemma~\ref{lemma:r} we have that
\[
\frac{\deg_M(V(\calX))}{\deg_{M^*}(V(\calX))} 
\geq \half + \inv{2\cmax} - \frac{\cmax(\cmax+1)}{2k}
~.
\]
Since $\deg_{M^*}(V(\calT)) = \sum_{\calX} \deg_{M^*}(V(\calX))$ and
$\deg_M(V(\calT)) = \sum_{\calX} \deg_M(V(\calX))$, it follows that
\[
\frac{\deg_M(V(\calT))}{\deg_{M^*}(V(\calT))}
\geq \half + \inv{2\cmax} - \frac{\cmax(\cmax+1)}{2k}
~,
\]
and thus
\(
\abs{M \cap A(\calT)}
\geq \paren{ \half + \inv{2\cmax} -
             \frac{\cmax(\cmax+1)}{2k}} \abs{M^* \cap A(\calT)}
%~,
\).
%and the lemma follows.
\end{proof}

By setting $k = \ceil{\cmax(\cmax+1)/2\eps}$, we get the following
result.

\begin{corollary}
\label{cor:local}
There exists a $(\half + \inv{2\cmax} - \eps)$-approximation algorithm
for unweighted \carpool, for every $\eps>0$.
\end{corollary}

In the Appendix~\ref{sec:tight} we show that our analysis is tight.