### 2017-02-15 src/in/tikz.tex
fix some refs

\usepackage{tikz}
\usetikzlibrary{
	arrows.meta,
	calc,
	graphs,
	patterns,
	positioning,
	shapes,
	decorations.pathmorphing,
	math,
}

% NODE
\tikzset{default node/.style={draw,circle,inner sep=0mm,minimum size=5mm,very thick,font=\small,black!70,}}

% LABELS
\tikzset{label/.style={draw=none,sloped,}}
\tikzset{label above/.style={label,midway,above=-1mm,}}
\tikzset{label below/.style={label,midway,below=-1mm,}}

% EDGES
\tikzset{e1/.style={cyan}}
\tikzset{e2/.style={green, loosely dash dot dot}}
\tikzset{e3/.style={blue, loosely dotted}}
\tikzset{e4/.style={loosely dashed}}
\tikzset{e5/.style={red, loosely dashed}}
\tikzset{e6/.style={orange, loosely dash dot}}

% LINES
\tikzset{m/.style={blue, dashed}}
\tikzset{outline/.style={loosely dashed, ultra thin}}

%DEFAULT STYLE
\tikzset{
	>=Triangle[], 
	on grid, 
	auto, 
	thick,
}