### 2017-02-17 src/abstract.tex
merge dror

The \carpool problem is a \emph{star packing} problem in directed
graphs.  Formally, given a directed graph $G = (V, A)$, a capacity
function $c: V \rightarrow \N$, and a weight function $w:
A \rightarrow \R^+$, a \emph{carpool matching} is a subset of arcs,
$M \subseteq A$, such that every $v \in V$ satisfies:%
\begin{inparaenum}[(i)]
\item $\din[M](v) \cdot \dout[M](v) = 0$,
\item $\din[M](v) \leq c(v)$, and 
\item $\dout[M](v) \leq 1$.
\end{inparaenum}
A vertex $v$ for which $\dout[M](v) = 1$ is a \emph{passenger}, and a
vertex for which $\dout[M](v) = 0$ is a \emph{driver} who has
$\din[M](v)$ passengers.  In the \carpool problem the goal is to find
a carpool matching $M$ of maximum total weight.
%
The problem arises when designing an online carpool service, such as
Zimride~\cite{zimride}, which tries to connect between users based on
a similarity function.  The problem is known to be NP-hard, even in
the unweighted and uncapacitated case.
%
The \gcp problem, is an extension of the \carpool where each vertex
represents an unsplittable group of passengers.  Formally, each vertex
$u \in V$ has a size $s(u) \in \N$, and the constraint $\din(v) \leq
c(v)$ is replaced with $\sum_{u:(u,v) \in M} s(u) \leq c(v)$.

We show that \carpool can be formulated as an unconstrained submodular
maximization problem, thus it admits a $\frac{1}{2}$-approximation
algorithm.  We show that the same formulation does not work for \gcp,
nevertheless, we present a local search $(\frac{1}{2}
- \eps)$-approximation algorithm for \gcp.
%
For the unweighted variant of both problems when the maximum possible
capacity, $\cmax$, is bounded by a constant, we provide a local search
$(\half + \inv{2\cmax} - \eps)$-approximation algorithm.  We also show
that the problem APX-hard, even if the maximum degree (and $\cmax$) is
at most $3$.