### 2017-10-29 src/main.tex
rm appendix

\documentclass[a4paper,UKenglish]{lipics-v2016}

\input{def}
\input{tikz}
\input{libs}

\renewcommand{\floatpagefraction}{0.8}

\sloppy

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{title}

\begin{document}

\maketitle

\input{outline}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage

\subparagraph*{Acknowledgements.}

We thank David Adjiashvili and Reuven Bar-Yehuda for helpful
discussions.

\bibliographystyle{plainurl}
\bibliography{main}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}